\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-05-23 doc/paper.Rmd
added high-resolution scan image

\autoref{fig:eer}
\autoref{tab:aucs}

@@

p1 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
  ggplot(aes( x = wo, y=eer, colour=method, shape=method)) +
  geom_point() +
  facet_wrap(~wv, labeller="label_both") + 
  theme_bw() +
  ylab("Equal Error rate (EER)") +
  xlab("Size of optimization window in pixels") +
  theme(legend.position = "bottom") + #c(0.25, 0.25)) +
  scale_colour_brewer(palette="Set1") +
  scale_y_continuous(limits=c(0, 0.25))

p2 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
  ggplot(aes( x = wo, y=AUC, colour = method, shape=method)) +
  geom_point(size=2) +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=5) +
  facet_wrap(~wv, labeller="label_both") + 
  theme_bw() +
  ylab("Area under the curve (AUC)") +
  xlab("Size of optimization window in pixels")+
  theme(legend.position = "bottom") + #c(0.75, 0.25)) +
  scale_colour_brewer(palette="Set1")
grid.arrange(p1, p2, nrow=1)
```

@@

\begin{table}[ht]
\centering
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{rrrlrrl}
  \hline
  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\ 
  \hline
90 & 0.068 & 0.330 & 0.850 (0.835, 0.865) & 0.075 & 0.242 & 0.877 (0.863, 0.890) \\ 
  120 & 0.065 & 0.309 & 0.864 (0.850, 0.878) & 0.072 & 0.236 & 0.890 (0.877, 0.903) \\ 
  130 & 0.065 & 0.317 & 0.863 (0.850, 0.877) & 0.072 & 0.224 & 0.898 (0.886, 0.910) \\ 
  140 & 0.065 & 0.304 & 0.873 (0.859, 0.886) & 0.071 & 0.229 & 0.902 (0.891, 0.914) \\ 
  150 & 0.067 & 0.326 & 0.863 (0.850, 0.877) & 0.072 & 0.225 & 0.904 (0.892, 0.916) \\ 
  180 & 0.061 & 0.326 & 0.877 (0.864, 0.890) & 0.070 & 0.229 & 0.907 (0.896, 0.919) \\ 
  210 & 0.064 & 0.342 & 0.865 (0.851, 0.878) & 0.066 & 0.231 & 0.906 (0.895, 0.918) \\ 
   \hline
\end{tabular}
\end{adjustbox}
\caption{\label{tab:aucs} Overview of results as shown in Figures \ref{fig:roc} and \ref{fig:eer}. FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$) or type I error rate, FNR is the false negative rate or type II eror rate. Area under the curve (AUC) is shown with confidence intervals \citep{delong}. CS2 significantly outperforms CS1 with respect to its predictive power. }
\end{table}

@@

\autoref{fig:type1} shows observed type I error rates across a range of optimization windows $w_o$. Generally, observed type I errors are higher than expected. Method CS1 shows in this instance slightly better performance than method CS2, but for both an increase in the size of the optimization window leads to a decrease in the observed type I error rates.

```{r type1, fig.width=7.75, fig.height=5.25, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$. The horizontal line in each facet indicates the nominal type I error rate."}
errorsCV$wv.method <- paste(errorsCV$method, errorsCV$wv, sep=" / ")
errorsCV %>% filter(wv %in% c(30,50)) %>%
  ggplot(aes(x = wo, y = actual, colour=factor(alpha), 
             linetype=method)) + 
  geom_hline(aes(yintercept=alpha), colour="grey30", size=.5) +
  geom_point(aes(shape=factor(wv.method)), size=2.5) +
  geom_smooth(aes(group=factor(method)), 
              size=.8, se=FALSE, method="lm", alpha=.9) +
  theme_bw() +
  ylab("Observed type I error rate") +
  xlab(expression("Window size for optimization, "~w[o])) +
  scale_colour_brewer(
    expression("Nominal type I error "~alpha), palette="Set2") +
  facet_wrap(~alpha, labeller="label_both", scales="free") +
  theme(legend.position="bottom") +
  scale_shape_manual(expression("Method / Validation window "~w[v]), values=c(1,2,16,17)) +
  scale_linetype_manual("Method", values=c(3,1), guide=FALSE) +
  guides(shape = guide_legend(ncol=2),
         col = guide_legend(ncol=2))

>>>>

\autoref{fig:eer} shows a comparison of the performance of the two methods CS1 and CS2 with respect to EER (eaual error rate) and AUC (area under the curve) corresponding to the ROC curves shown in \autoref{fig:roc}. Equal error rates are reduced using method CS2, while area under the curve significantly increases (at a significance level $\alpha$ of 0.05) compared to method CS1.

The results from Figures \ref{fig:roc} and \ref{fig:eer} are summarized in numbers in \autoref{tab:aucs}. Equal error rates (EER), rates for false positives (FPR) and false negatives (FNR) are shown side by side with the area under the curve (AUR) for both methods for a set of different optimization windows $w_o$ and a validation window $w_v$ of 30 pixels. The rate of false positive same-source identifications is equal to the statistical type I error rate, which is set to $\alpha = 0.05$ for this example. The rate of false negatives are missed same-source markings. This rate is also known as the type II error rate.   
Area under the curve (AUC) is shown with confidence intervals as given by \citet{delong}. CS2 significantly outperforms CS1 with respect to its predictive power in most situations.

@@

p1 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
  ggplot(aes( x = wo, y=eer, colour=method, shape=method)) +
  geom_point(size=2) +
  facet_wrap(~wv, labeller = label_bquote(w[v] == .(wv))) + 
  theme_bw() +
  ylab("Equal Error rate (EER)") +
  xlab("Size of optimization window in pixels") +
  theme(legend.position = "bottom") + #c(0.25, 0.25)) +
  scale_colour_brewer("Method", palette="Set1") +
  scale_shape("Method") +
  scale_y_continuous(limits=c(0, 0.25))

p2 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
  ggplot(aes( x = wo, y=AUC, colour = method, shape=method)) +
  geom_point(size=2) +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=5) +
  facet_wrap(~wv, labeller = label_bquote(w[v] == .(wv))) + 
  theme_bw() +
  ylab("Area under the curve (AUC)") +
  xlab("Size of optimization window in pixels")+
  theme(legend.position = "bottom") + #c(0.75, 0.25)) +
  scale_colour_brewer("Method", palette="Set1") +
  scale_shape("Method") 
grid.arrange(p1, p2, nrow=1)
```

@@

\begin{table}[ht]
\centering
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{rrrlrrl}
  \hline
  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\ 
  \hline
90 & 0.068 & 0.330 & 0.850 (0.835, 0.865) & 0.075 & 0.242 & 0.877 (0.863, 0.890) \\ 
  120 & 0.065 & 0.309 & 0.864 (0.850, 0.878) & 0.072 & 0.236 & 0.890 (0.877, 0.903) \\ 
  130 & 0.065 & 0.317 & 0.863 (0.850, 0.877) & 0.072 & 0.224 & 0.898 (0.886, 0.910) \\ 
  140 & 0.065 & 0.304 & 0.873 (0.859, 0.886) & 0.071 & 0.229 & 0.902 (0.891, 0.914) \\ 
  150 & 0.067 & 0.326 & 0.863 (0.850, 0.877) & 0.072 & 0.225 & 0.904 (0.892, 0.916) \\ 
  180 & 0.061 & 0.326 & 0.877 (0.864, 0.890) & 0.070 & 0.229 & 0.907 (0.896, 0.919) \\ 
  210 & 0.064 & 0.342 & 0.865 (0.851, 0.878) & 0.066 & 0.231 & 0.906 (0.895, 0.918) \\ 
   \hline
\end{tabular}
\end{adjustbox}
\caption{\label{tab:aucs} Overview of results as shown in Figures \ref{fig:roc} and \ref{fig:eer}. FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$) or type I error rate, FNR is the false negative rate or type II eror rate. Area under the curve (AUC) is shown with confidence intervals. }
\end{table}

@@

\autoref{fig:type1} shows observed type I error rates (FPR) across a range of optimization windows $w_o$. Generally, observed type I errors are higher than expected. Method CS1 shows in this instance slightly better performance than method CS2, but for both an increase in the size of the optimization window leads to a decrease in the observed type I error rates.

```{r type1, fig.width=7.75, fig.height=5.25, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$. The horizontal line in each facet indicates the nominal type I error rate."}
errorsCV$wv.method <- paste(errorsCV$method, errorsCV$wv, sep=" / ")
errorsCV %>% filter(wv %in% c(30,50)) %>%
  ggplot(aes(x = wo, y = actual, colour=factor(alpha), 
             linetype=method)) + 
  geom_hline(aes(yintercept=alpha), colour="grey30", size=.5) +
  geom_point(aes(shape=factor(wv.method)), size=2.5) +
  geom_smooth(aes(group=factor(method)), 
              size=.8, se=FALSE, method="lm", alpha=.9) +
  theme_bw() +
  ylab(expression("Observed type I error rate "~hat(alpha))) +
  xlab(expression("Window size for optimization "~w[o])) +
  scale_colour_brewer(
    expression("Nominal type I error "~alpha), palette="Set2") +
  facet_wrap(~alpha, labeller=label_bquote(alpha == .(alpha)), scales="free") +
  theme(legend.position="bottom") +
  scale_shape_manual(expression("Method / Validation window "~w[v]), values=c(1,2,16,17)) +
  scale_linetype_manual("Method", values=c(3,1), guide=FALSE) +
  guides(shape = guide_legend(ncol=2),
         col = guide_legend(ncol=2))

## High resolution Hamby 44 scans

```{r h44, fig.width=7.75, fig.height=5.25, out.width='\\textwidth', fig.cap="AUC and EER for methods CS1 and CS@ based on high resolution scans of Hamby 44. "}
if (!file.exists("../data/h44.rds")) {
files <- dir("../data/hamby44/", pattern="rds", recursive=TRUE)

h44 <- data.frame()
for (file in files) {
  tmp <- readRDS(file.path("../data/hamby44", file))
  tmp$source <- file
  tmp <- tmp %>% mutate(
    match = replace(match, is.na(match), FALSE)
  )
  h44 <- rbind(h44, tmp)
}
  saveRDS(h44, "../data/h44.rds")
} else {
  readRDS("../data/h44.rds")
}

h44 <- h44 %>% separate(source, into=c("method", "rest"), sep="/") %>%
  dplyr::select(-rest)

h44.summ <- h44 %>% group_by(wo, wv, method, coarseness) %>%
  mutate(
    fails = sum(is.na(p_value))/length(p_value)
  ) %>% filter(!is.na(p_value)) %>%
  summarize(
    AUC = AUC(1-p_value, match),
    EER = EER(1-p_value, match),
  #  ci = list(AUC_confint(1-p_value, match)),
    fails = fails[1],
    delong = list(pROC::ci.auc(match, 1-p_value))
  )

h44.summ <- h44.summ %>% mutate(
#  lower = ci %>% purrr::map_dbl(.f = function(x) x[[1]][1]),
#  upper = ci %>% purrr::map_dbl(.f = function(x) x[[1]][2]),
  low_delong = delong %>% purrr::map_dbl(.f = function(x) x[1]),
  upp_delong = delong %>% purrr::map_dbl(.f = function(x) x[3]),
  Method = toupper(method)
)

h44.summ %>% gather(type, value, AUC, EER) %>% 
  filter(wv %in% c(75, 125), coarseness==0.125,
         between(wo, 210, 390)) %>% 
  ggplot(aes(x = wo, y = value, colour=Method, shape=factor(wv)))  +
  geom_point(size=2.5) +
#  geom_point(aes(y = lower), size=1) +
#  geom_point(aes(y = upper), size=1) +
  facet_wrap(~type, scales="free") +
  theme_bw() +
  scale_colour_brewer(palette="Set1", guide = guide_legend(ncol = 1)) +
  geom_smooth(aes(linetype = factor(wv), colour=Method), se=FALSE, span=2) +
  xlab(expression("Size of optimization window "~w[o]~" in pixels")) +
  ylab("Value") +
  theme(legend.position = "bottom", 
        legend.key.width = unit(2,"line")) +
  scale_shape(expression("Size of validation window "~w[v]~" in pixels"), guide = guide_legend(ncol=1))+
  scale_linetype(expression("Size of validation window "~w[v]~" in pixels"), guide = guide_legend(ncol=1))

### 2018-05-23 doc/paper.tex
added high-resolution scan image

\autoref{fig:eer}
\autoref{tab:aucs}

@@

{\textcolor{orange}{
We started out trying to assess the suitability of the Chumbley Score's suitability for matching striae on bullet lands. Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable. 
Optimized window sizes were found based on cross-validation to minimize type 2 error rates. 
Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of XXX for an optimized window size of YYY - which is considerably higher than the error rates achieved on matching toolmarks. 
Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
XXX
The algorithm CS2 introduced here is constructed in a way that reduces the number of failures by a magnitude of about XXX. While reducing the failure rate, the algorithm also shows an increase in the power of the test. Type II error rates reach a minimum of XXX for an optimized window size of YYY. While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the -much larger- toolmarks.
However, bullets usually have multiple lands - in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet. We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
}}

>>>>

\autoref{fig:eer} shows a comparison of the performance of the two
methods CS1 and CS2 with respect to EER (eaual error rate) and AUC (area
under the curve) corresponding to the ROC curves shown in
\autoref{fig:roc}. Equal error rates are reduced using method CS2, while
area under the curve significantly increases (at a significance level
\(\alpha\) of 0.05) compared to method CS1.

The results from Figures \ref{fig:roc} and \ref{fig:eer} are summarized
in numbers in \autoref{tab:aucs}. Equal error rates (EER), rates for
false positives (FPR) and false negatives (FNR) are shown side by side
with the area under the curve (AUR) for both methods for a set of
different optimization windows \(w_o\) and a validation window \(w_v\)
of 30 pixels. The rate of false positive same-source identifications is
equal to the statistical type I error rate, which is set to
\(\alpha = 0.05\) for this example. The rate of false negatives are
missed same-source markings. This rate is also known as the type II
error rate.\\
Area under the curve (AUC) is shown with confidence intervals as given
by \citet{delong}. CS2 significantly outperforms CS1 with respect to its
predictive power in most situations.

@@

{\textcolor{orange}{
We started out trying to assess the suitability of the Chumbley Score's for matching striae on bullet lands. Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable. 
Optimized window sizes were found based on cross-validation to minimize type 2 error rates. 
Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of XXX for an optimized window size of YYY - which is considerably higher than the error rates achieved on matching toolmarks. 
Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
XXX
The algorithm CS2 introduced here is constructed in a way that reduces the number of failures by a magnitude of about XXX. While reducing the failure rate, the algorithm also shows an increase in the power of the test. Type II error rates reach a minimum of XXX for an optimized window size of YYY. While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the -much larger- toolmarks.
However, bullets usually have multiple lands - in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet. We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
}}
